---
title: "Tape Intervention R Script"
author: "Cas Huisman"
format: 
  html:
    code-fold: true
    code-summary: "Show R script"
params:
  path: "C:/Users/huism080/OneDrive - Wageningen University & Research/Research/Study 2/tapeStudy/data/" #specify the folder where the anonymised data is located
---
```{r}
#| echo: false
#| include: false
#| label: libraries

library(table1) # sample characteristics
library(vtable) # summary table
library(ggplot2) # visualisation
library(dplyr) # data tidying
library(AER) # tobit models
library(lme4) # mixed regression models
library(nlme)
library(broom) # tidy tables
library(broom.mixed) # broom for mixed models
library(flextable) # nice tables
library(emmeans) # emmeans for post-hoc
library(tidyr)
library(showtext)

showtext_auto()
```

```{r}
#| label: read data
#| echo: false

t0 <- read.csv(url("https://raw.githubusercontent.com/CasHsmn/R_tape/main/data/t0_anon.csv"))
t1 <- read.csv(url("https://raw.githubusercontent.com/CasHsmn/R_tape/main/data/t1_anon.csv"))
t2 <- read.csv(url("https://raw.githubusercontent.com/CasHsmn/R_tape/main/data/t2_anon.csv"))
tall3 <- read.csv(url("https://raw.githubusercontent.com/CasHsmn/R_tape/main/data/tall_anon.csv"))

tall <- tall3[tall3$time != 3,]

```

# Sample Characteristics
```{r}
#| label: sample characteristics
#| echo: false

# Sample descriptives; Label sample variables and create table 1
t0$gender[t0$gender == 4] <- 3
t0$gender <- factor(t0$gender, levels = c(1,2,3), labels = c("Female", "Male", "Other"))
t0$condition <- factor(t0$condition, levels = c("intervention", "control"), labels = c("intervention", "control"))
tall$condition <- factor(tall$condition)
label(t0$gender) <- "Gender"
label(t0$age) <- "Age"
label(t0$hh_size) <- "Household size"
label(t0$fw_g) <- "Baseline waste"
table1(~ gender + age + hh_size + fw_g | condition, data = t0)
```

# Descriptives
```{r}
#| label: descriptives
#| echo: false
#| warning: false

st(tall, vars = c("pastweek_1", "pastweek_2", "pastweek_3", "stock_freq_1", "hs_stock", "meal_freq._1", "hs_meal", "snack_freq_1", "hs_snack", "share_freq_1", "hs_share"), labels = c("Days at home", "Days ate at home", "Days prepared food", "Check if near expiry", "Habit: Check if near expired", "Use near expired food in meal", "Habit: Use near expired food in meal", "Snack near expired", "Habit: Snack near expired", "Share near expired", "Habit: Share near expired"), summ = c("mean(x)", "sd(x)", "median(x)"), out = 'kable', title = "Overall variable descriptives", col.align = 'center')
```

```{r}
#| label: barplots of food waste by condition and time
#| tbl-cap: Bar plot
#| ncol: 2
#| fig-cap-location: top
#| fig-cap: Bar plots
#| fig-subcap: 
#| - Food waste over time by condition with 95% CI
#| - The log of food waste over time by condition with 95% CI
#| warning: false


tall$time <- factor(tall$time, levels = c(0,1,2), labels = c("pre", "post", "follow-up"))

tall %>% 
  group_by(condition, time) %>% 
  summarise(
    n=n(),
    mean=mean(fw_g),
    sd=sd(fw_g)
  ) %>% 
  mutate(se = sd/sqrt(n)) %>% 
  mutate(ci=se*qt((1-0.05)/2 + .5, n-1)) %>% 
  ggplot(aes(x = time, y = mean, fill = condition)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  labs(title = "Food waste (g) over time by condition",
       x = "Time",
       y = "Food Waste (g)",
       fill = "Condition") +
  theme(plot.title = element_text(size =16, family = "Arvo"),
        axis.title = element_text(size = 14, family = "TT Arial"),
        axis.text = element_text(size = 12, family = "TT Arial"),
        legend.title = element_text(size = 14, family = "TT Arial"),
        legend.text = element_text(size = 12, family = "TT Arial"),
        axis.line = element_line(colour = "darkgray", size = 1, linetype = "solid"),
        panel.background = element_rect(fill = "transparent"),
        panel.grid.major = element_line(size = .05, linetype = "solid", colour = "grey")) +
  scale_y_continuous(expand = c(0, 0), limits = c(0,450)) +
  scale_fill_brewer(palette = "Pastel2")

tall <- mutate(tall, fw_g_log = log(fw_g + 1))

tall %>% 
  group_by(condition, time) %>% 
  summarise(
    n=n(),
    mean=mean(fw_g_log),
    sd=sd(fw_g_log)
  ) %>% 
  mutate(se = sd/sqrt(n)) %>% 
  mutate(ci=se*qt((1-0.05)/2 + .5, n-1)) %>% 
  ggplot(aes(x = time, y = mean, fill = condition)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  labs(title = "Food waste over time by condition",
       x = "Time",
       y = "Food Waste (log(g))",
       fill = "Condition") +
  theme(plot.title = element_text(size =18, family = "Arvo"),
        axis.title = element_text(size = 16, family = "TT Arial"),
        axis.text = element_text(size = 14, family = "TT Arial"),
        legend.title = element_text(size = 16, family = "TT Arial"),
        legend.text = element_text(size = 14, family = "TT Arial"),
        axis.line = element_line(colour = "darkgray", size = 1, linetype = "solid"),
        panel.background = element_rect(fill = "transparent"),
        panel.grid.major = element_line(size = .05, linetype = "solid", colour = "grey")) +
  scale_y_continuous(expand = c(0, 0), limits = c(0,NA)) +
  scale_fill_brewer(palette = "Pastel2")

```

## Food waste distribution

The distribution of food waste in the sample is as follows:

As the histograms below show, food waste is very right skewed with many 0s. The distribution is better after log transformation.
```{r}

ggplot(tall, aes(x = fw_g)) +
  geom_histogram(binwidth = 20)+
  labs(title = "Histogram of food waste (g)")

ggplot(tall, aes(x = fw_g_log)) +
  geom_histogram(binwidth = .2)  +
  labs(title = "Histogram of log transformed food waste")
```
# Inferential statistics


No significant effect for age, gender and household size on food waste. They are not included as covariates.
```{r}
#| label: hypothesis-testing
#| tbl-cap: Regression of demographics predicting FW
#| tbl-cap-location: top

tall <- tall %>%
  group_by(id) %>%
  mutate(age = ifelse(is.na(age), first(age), age),
         gender = ifelse(is.na(gender), first(gender), gender)) %>%
  ungroup()

tall$gender[tall$gender == 4] <- 3
tall$gender <- factor(tall$gender, levels = c(1,2,3), labels = c("Female", "Male", "Other"))
lm_fw_1 <- lmer(fw_g_log ~ age + gender + hh_size + (1|id), data = tall)

flextable(tidy(lm_fw_1)) %>% colformat_double(digits = 2)

lm_fw_0 <- lme(fw_g ~ pastweek_1 + pastweek_2 + pastweek_3, random = ~1 | id, data = tall) # mixed model for the control variables of days home/cooked etc.
```



```{r}
#| tbl-cap: Mixed model regression of food waste predicted by condition*time
#| tbl-cap-location: top
lm_fw_2 <- lme(fw_g_log ~ condition * time, random = ~1 | id, data = tall)
flextable(tidy(lm_fw_2)) %>% colformat_double(digits = 2) %>% set_caption(caption = "Coefficients for mixed model predicting log transformed fw(g)")

nlme_fw <- lme(fw_g ~ condition * time, random = ~1 | id, data = tall)
flextable(tidy(nlme_fw)) %>% colformat_double(digits = 2) %>% set_caption(caption = "Coefficients for mixed model predicting fw(g)")

```

# Contrasts
The test shows a significant interaction of condition over time, so I will calculate emmeans and contrasts.
```{r}
fw_emmeans <- emmeans(nlme_fw, specs = ~ condition | time)
contrasts <- contrast(fw_emmeans, method = "pairwise", adjust = "tukey")
contrasts_time <- contrast(fw_emmeans, method = "pairwise", by = "condition")
flextable(tidy(contrasts_time)) %>% colformat_double(digits = 3)

flextable(tidy(contrasts)) %>% colformat_double(digits = 3)

fw_emmeans_df <- as.data.frame(fw_emmeans)

ggplot(fw_emmeans_df, aes(x = time, y = emmean, color = condition, group = condition)) +
  geom_point(size = 3) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of fw_g_log by Condition and Time",
       x = "Time",
       y = "Log Transformed Food Waste (fw_g_log)",
       color = "Condition") +
  theme_minimal()
```

In the contrasts it shows that:

* There is a significant drop in food waste for the intervention group from baseline to post-measure, but not in the control group
* There is a marginally significant drop in food waste for the intervention group from post to follow-up
* at baseline, there is no difference between control and intervention
* at post and followup, people in the intervention group waste significantly less than those in the control group


# Habit analysis
In this part I will first look if habit strength of the four behaviours can be predicted by condition*time

## Habit of checking which food is near expiry
```{r}


lme_stockhs <- lme(hs_stock ~ condition*time, random = ~1|id, data = tall3)
flextable(tidy(lme_stockhs)) %>% colformat_double(digits = 2)

emmeans_stockhs <- emmeans(lme_stockhs, specs = ~ condition | time)
contrasts_stockhs <- contrast(emmeans_stockhs, method = "pairwise", adjust = "tukey")
contrasts_time_stockhs <- contrast(emmeans_stockhs, method = "pairwise", by = "condition")
flextable(tidy(contrasts_time_stockhs)) %>% colformat_double(digits = 3)

flextable(tidy(contrasts_stockhs)) %>% colformat_double(digits = 3)

emmeans_stock_df <- as.data.frame(emmeans_stockhs)

ggplot(emmeans_stock_df, aes(x = time, y = emmean, color = condition, group = condition)) +
  geom_point(size = 3) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of habit strength of stock checking by Condition and Time",
       x = "Time",
       y = "Habit strength",
       color = "Condition") +
  theme_minimal()
```
## Habit of letting housemates know they can eat food which is near expired
```{r}

lme_sharehs <- lme(hs_share ~ condition*time, random = ~1|id, data = tall3)
flextable(tidy(lme_sharehs)) %>% colformat_double(digits = 2)

emmeans_sharehs <- emmeans(lme_sharehs, specs = ~ condition | time)
contrasts_sharehs <- contrast(emmeans_sharehs, method = "pairwise", adjust = "tukey")
contrasts_time_sharehs <- contrast(emmeans_sharehs, method = "pairwise", by = "condition")
flextable(tidy(contrasts_time_sharehs)) %>% colformat_double(digits = 3)

flextable(tidy(contrasts_sharehs)) %>% colformat_double(digits = 3)

emmeans_share_df <- as.data.frame(emmeans_sharehs)

ggplot(emmeans_share_df, aes(x = time, y = emmean, color = condition, group = condition)) +
  geom_point(size = 3) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of habit strength of share checking by Condition and Time",
       x = "Time",
       y = "Habit strength",
       color = "Condition") +
  theme_minimal()
```

## Habit of using near expired foods when preparing meals
```{r}

lme_mealhs <- lme(hs_meal ~ condition*time, random = ~1|id, data = tall3)
flextable(tidy(lme_mealhs)) %>% colformat_double(digits = 2)

emmeans_mealhs <- emmeans(lme_mealhs, specs = ~ condition | time)
contrasts_mealhs <- contrast(emmeans_mealhs, method = "pairwise", adjust = "tukey")
contrasts_time_mealhs <- contrast(emmeans_mealhs, method = "pairwise", by = "condition")
flextable(tidy(contrasts_time_mealhs)) %>% colformat_double(digits = 3)

flextable(tidy(contrasts_mealhs)) %>% colformat_double(digits = 3)

emmeans_meal_df <- as.data.frame(emmeans_mealhs)

ggplot(emmeans_meal_df, aes(x = time, y = emmean, color = condition, group = condition)) +
  geom_point(size = 3) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of habit strength of meal checking by Condition and Time",
       x = "Time",
       y = "Habit strength",
       color = "Condition") +
  theme_minimal()
```
## Habit of picking something which is near expired as a snack
```{r}

lme_snackhs <- lme(hs_snack ~ condition*time, random = ~1|id, data = tall3)
flextable(tidy(lme_snackhs)) %>% colformat_double(digits = 2)

emmeans_snackhs <- emmeans(lme_snackhs, specs = ~ condition | time)
contrasts_snackhs <- contrast(emmeans_snackhs, method = "pairwise", adjust = "tukey")
contrasts_time_snackhs <- contrast(emmeans_snackhs, method = "pairwise", by = "condition")
flextable(tidy(contrasts_time_snackhs)) %>% colformat_double(digits = 3)

flextable(tidy(contrasts_snackhs)) %>% colformat_double(digits = 3)

emmeans_snack_df <- as.data.frame(emmeans_snackhs)

ggplot(emmeans_snack_df, aes(x = time, y = emmean, color = condition, group = condition)) +
  geom_point(size = 3) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of habit strength of snack checking by Condition and Time",
       x = "Time",
       y = "Habit strength",
       color = "Condition") +
  theme_minimal()
```

