---
title: "YRM Data Overview"
author: "Cas Huisman"
format: 
  html:
    code-fold: true
    code-summary: "Show R script"
params:
  path: "C:/Users/huism080/OneDrive - Wageningen University & Research/Research/Study 2/tapeStudy/data/" #specify the folder where the anonymised data is located
---

```{r}
#| echo: false
#| include: false
#| label: libraries

library(table1) # sample characteristics
library(vtable) # summary table
library(ggplot2) # visualisation
library(dplyr) # data tidying
library(AER) # tobit models
library(lme4) # mixed regression models
library(nlme)
library(broom) # tidy tables
library(broom.mixed) # broom for mixed models
library(flextable) # nice tables
library(emmeans) # emmeans for post-hoc
library(tidyr)
library(showtext)
library(rstatix)
library(car)
library(tibble)

showtext_auto()
```

```{r}
#| label: read data
#| echo: false

t0 <- read.csv(url("https://raw.githubusercontent.com/CasHsmn/R_tape/main/data/t0_anon.csv"))
t1 <- read.csv(url("https://raw.githubusercontent.com/CasHsmn/R_tape/main/data/t1_anon.csv"))
t2 <- read.csv(url("https://raw.githubusercontent.com/CasHsmn/R_tape/main/data/t2_anon.csv"))
tall3 <- read.csv(url("https://raw.githubusercontent.com/CasHsmn/R_tape/main/data/tall_anon.csv"))
tall3$condition <- factor(tall3$condition)
tall3$time <- factor(tall3$time, levels = c(0,1,2,3), labels = c("pre", "post", "post2", "post3"))
tall <- tall3[tall3$time != "post3",]

```

# Sample Characteristics
```{r}
#| label: sample characteristics
#| echo: false

# Sample descriptives; Label sample variables and create table 1
t0$gender[t0$gender == 4] <- 3
t0$gender <- factor(t0$gender, levels = c(1,2,3), labels = c("Female", "Male", "Other"))
t0$condition <- factor(t0$condition, levels = c("intervention", "control"), labels = c("intervention", "control"))
tall$condition <- factor(tall$condition)

label(t0$gender) <- "Gender"
label(t0$age) <- "Age"
label(t0$hh_size) <- "Household size"
label(t0$fw_g) <- "Baseline waste"
table1(~ gender + age + hh_size + fw_g | condition, data = t0)
```

# Descriptives
It looks like food waste is going down for the intervention group but not the control group.
```{r}
#| label: barplots of food waste by condition and time
#| tbl-cap: Bar plot
#| ncol: 2
#| fig-cap-location: top
#| fig-cap: Bar plots
#| fig-subcap: 
#| - Food waste over time by condition with 95% CI
#| - The log of food waste over time by condition with 95% CI
#| warning: false


tall$time <- factor(tall$time)
tall3$time <- factor(tall3$time, levels = c(0,1,2,3), labels = c("pre", "post", "follow-up","final"))


tall %>% 
  group_by(condition, time) %>% 
  summarise(
    n=n(),
    mean=mean(fw_g),
    sd=sd(fw_g)
  ) %>% 
  mutate(se = sd/sqrt(n)) %>% 
  mutate(ci=se*qt((1-0.05)/2 + .5, n-1)) %>% 
  ggplot(aes(x = time, y = mean, fill = condition)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  labs(title = "Food waste (g) over time by condition",
       x = "Time",
       y = "Food Waste (g)",
       fill = "Condition") +
  theme(plot.title = element_text(size =16, family = "Arvo"),
        axis.title = element_text(size = 14, family = "TT Arial"),
        axis.text = element_text(size = 12, family = "TT Arial"),
        legend.title = element_text(size = 14, family = "TT Arial"),
        legend.text = element_text(size = 12, family = "TT Arial"),
        axis.line = element_line(colour = "darkgray", size = 1, linetype = "solid"),
        panel.background = element_rect(fill = "transparent"),
        panel.grid.major = element_line(size = .05, linetype = "solid", colour = "grey")) +
  scale_y_continuous(expand = c(0, 0), limits = c(0,450)) +
  scale_fill_brewer(palette = "Pastel2")
```

But food waste always has a distribution with very many 0s and right-skew. However, the difference scores of food waste are normally distributed.
```{r}

ggplot(tall, aes(x = fw_g)) +
  geom_histogram(binwidth = 20)+
  labs(title = "Histogram of food waste (g)")

ggplot(tall, aes(x = fw_g_diff)) +
  geom_histogram(binwidth = 40)+
  labs(title = "Histogram of food waste (g) difference scores")
```
# Inferential statistics
if I just continue the analysis as is, using a mixed model, I get the following results.

A significant interaction for time*condition. And a significant decrease in food waste for the intervention group from pretest to the second post test.
```{r}
fw_lmer <- lmer(fw_g ~ condition*time + (1|id), data = tall)
flextable(Anova(fw_lmer, test.statistic = "F") %>% rownames_to_column(var = ".")) %>% colformat_double(digits = 3)

fw_lmer_em <- emmeans(fw_lmer, specs = ~ condition | time)
cont_time <- contrast(fw_lmer_em, method = "pairwise", by = "condition")

flextable(tidy(cont_time)) %>% colformat_double(digits = 3)

fw_lmer_em_df <- as.data.frame(fw_lmer_em)

ggplot(fw_lmer_em_df, aes(x = time, y = emmean, color = condition, group = condition)) +
  geom_point(size = 3) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = emmean - SE, ymax = emmean + SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of fw_g by Condition and Time",
       x = "Time",
       y = "Food waste (g)",
       color = "Condition") +
  theme_minimal()
```
The main results for ANOVA are a bit different and these again differ from results of a RM ANOVA in SPSS. This one also has a significant interaction, but it is stronger.
```{r}
fw_aov <- anova_test(dv = fw_g, wid = id, within = time, between = condition, data = tall)
flextable(fw_aov$ANOVA) %>% colformat_double(digits = 3)
```